Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-06-06 PAPER.tex
Wed Jun  6 14:45:35 EDT 2018

The resulting eigenspectrum as compared to the green curve (FG + EoR) in Figure \ref{fig:toy_sigloss2} is shown in Figure 
\ref{fig:toy_sigloss15}. With 
fringe-rate filtering resulting in fewer independent modes, it becomes more difficult for the empirical covariance to estimate the 
true covariance matrix of the fringe-rate filtered data. This can be quantified by evaluating a convergence metric $\varepsilon(\Chat)$ for the empirical covariance, which we define as
\begin{equation}
\label{eq:converge}
\varepsilon (\Chat) \equiv \sqrt{\frac{\sum_{ij} (\widehat{C}_{ij} - {C}_{ij})^2}{\sum_{ij} {C}_{ij}^2}},
\end{equation}
where $\C$ is the true covariance matrix. To compute this metric, we draw different numbers of realizations (different draws of Gaussian noise) of our toy model EoR measurement, $\textbf{x}_{\rm EoR}$, and take their ensemble average. We then compare this to the ``true' covariance, which in our simulation is set to be the empirical covariance after a large number ($500$) of realizations. As shown in Figure \ref{fig:toy_sigloss16}, we perform this computation for a range of total independent ensemble realizations (horizontal axis) and the number of independent samples in the data following fringe-rate filtering (different colors). With more independent time samples (i.e. more realizations) in the data, one converges to the true fringe-rate filtered covariance more quickly.

@@

Just as important as the eigenvalues are the eigenvectors of our empirical covariances. In general, different eigenvectors converge to their true forms at different rates. This is illustrated by Figure \ref{fig:toy_sigloss17}, which shows the convergence of eigenvectors in an empirical estimate of a covariance matrix whose true form is a diagonal matrix with eigenvalues spanning four orders of magnitude. We use as a convergence metric $\varepsilon(\widehat{\textbf{v}})$ for the empirical eigenvectors $\widehat{\textbf{v}}$, defined as:
\begin{equation}
\label{eq:converge_eig}
\varepsilon (\widehat{\textbf{v}}) \equiv \sqrt{\sum_{i}^{N_{f}}|\textbf{v}-\widehat{\textbf{v}}|_{i}^2},
\end{equation}
where $N_{f}$ is the number of frequencies ($20$) in the mock data. The eigenmode convergence curves  in Figure \ref{fig:toy_sigloss17} are ranked ordered by eigenvalue, such that ``Eigenmode \#0" illustrates the convergence of the eigenvector with the largest eigenvalue, ``Eigenmode \#1" for the second largest eigenvalue, and so on. One sees that the highest-valued eigenmodes converge to their true eigenvectors more quickly. With only a small number of realizations, these empirically estimated modes already retain little correlation with the specific realizations of data that were used to form the empirical covariance. As we will see in the next section, using only these eigenmodes, which are less coupled to data realizations, minimizes signal loss. In contrast, the lower-valued eigenmodes retain more memory of the data realizations, which leads to correlations that induce signal loss. Said differently, a steep covariance eigenspectrum can be especially dangerous because the eigenmodes that converge the slowest are typically also EoR-dominated and are therefore susceptible to the most loss.  %\acl{Think about whether we need to do the Fourier space version as well.}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.31\textwidth]{plots/toy_sigloss15.pdf}
	\caption{Eigenspectrum of $\widehat{\textbf{C}}_{\rm FG+EoR}$, in the case of no fringe-rate filtering (solid green) and with fringe-rate filtering (dashed green). The dashed, steep eigenspectrum has a greater risk of signal loss because its lowest-valued eigenmodes are both more strongly coupled to the data (than those of the solid eigenspectrum) and, in this example, are also EoR-dominated.}
	\label{fig:toy_sigloss15}
\end{figure}

@@

Using our toy model, we have seen that a sensitivity-driven analysis technique like fringe-rate filtering has trade-offs of signal 
loss and noisier estimates when using data-estimated covariance matrices. Longer integrations increase sensitivity but reduce 
the number of independent samples, resulting in poorly characterized, steep eigenspectra that can overfit signal greatly. We 
note that a fringe-rate filter does have a range of benefits, many described in \citet{parsons_et_al2016}, so it can still be 
advantageous to use one despite the trade-offs.

@@

In Section \ref{sec:toymodel} we showed one example of how altering $\widehat{\textbf{C}}$ can 
make the difference between nearly zero and some signal loss. We will now use our toy model to describe several other ways to tailor $\widehat{\textbf{C}}$ 
in order to minimize signal loss. We choose four independent regularization methods to highlight in this section, which have 
been chosen due to their simplicity in implementation and straightforward interpretations. We illustrate the resulting power 
spectra and eigenspectra for the different cases in Figures \ref{fig:toy_sigloss8} and \ref{fig:toy_sigloss14}. These examples are not meant to be taken as suggested analysis methods but rather as illustrative cases.

As a first test, we model the covariance matrix of EoR as a proof of concept that if perfect models are known, signal loss can be 
avoided. We know that our simulated EoR signal should have a covariance matrix that mimics the identity matrix, with its 
variance encoded along the diagonal. We model $\textbf{C}_{\rm EoR}$ as such (i.e. the identity), instead of computing it based on $\textbf{x}
_{\rm EoR}$ itself. Next, we add $\textbf{C}_{\rm EoR} + \widehat{\textbf{C}}_{\rm FG}$ (where $\widehat{\textbf{C}}_{\rm FG} = \langle\textbf{x}_{\rm FG}
\textbf{x}_{\rm FG}^{\dagger}\rangle_{t}$) to obtain a final $\widehat{\textbf{C}}$ to use in weighting. In Figure \ref{fig:toy_sigloss8} (upper 
left), we see that there is negligible signal loss. This is because by modeling $\textbf{C}_{\rm EoR}$, we avoid over-fitting EoR fluctuations in the data that our model doesn't know about (but an empirically derived $\widehat{\textbf{C}}_{\rm EoR}$ would). This is also shown by comparing the (steeper) green and (flatter) red curves in Figure \ref{fig:toy_sigloss14}. In practice such a weighting option is not feasible, as it is difficult to model $\textbf{C}_{\rm EoR}$, and $\widehat{\textbf{C}}_{\rm FG}$ is unknown because we don't know how to separate out the foregrounds from the EoR in our data.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss14.pdf}
	\caption{We compare the eigenspectrum of an empirically calculated $\widehat{\textbf{C}}$ (green) to that of four alternate 
weighting options, including modeling the covariance matrix of EoR (red), regularizing $\widehat{\textbf{C}}$ by adding an identity 
matrix to it (blue), using only the first three eigenmodes of $\widehat{\textbf{C}}$ (yellow), and multiplying an identity matrix with $
\widehat{\textbf{C}}$ (magenta). All eigenspectra (except the green) are relatively flat and don't exhibit signal loss. All were 
computed for the `fringe-rate filtered' (time-averaged) toy model case presented in Section \ref{sec:toymodel_frf}.}
	\label{fig:toy_sigloss14}
\end{figure}

The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\widehat{\textbf{C}} \equiv 
\widehat{\textbf{C}} + \gamma\textbf{I}$, where $\gamma = 5$ (an arbitrary strength 
of $\textbf{I}$ for the purpose of this toy model). By adding the identity matrix, element-wise, we are weighting the diagonal 
elements of the estimated covariance matrix more heavily than those off-diagonal. Since the identity component does not know anything about the data realization, it alters the covariance to be less coupled to the data. Although there is negligible signal loss using this regularization, the small green peak at the third $k$-mode represents residual foregrounds that still exist since the shapes encoded in the off-diagonal frequency correlations of the covariance matrix were deemed not as prominent as the diagonal elements using this weighting scheme.

>>>>

% JEA - Death to eigenspectra!
%The resulting eigenspectrum as compared to the green curve (FG + EoR) in Figure \ref{fig:toy_sigloss2} is shown in Figure \ref{fig:toy_sigloss15}. 
With fringe-rate filtering resulting in fewer independent modes, it becomes more difficult for the empirical covariance to estimate the true covariance matrix of the fringe-rate filtered data. This can be quantified by evaluating a convergence metric $\varepsilon(\Chat)$ for the empirical covariance, which we define as
\begin{equation}
\label{eq:converge}
\varepsilon (\Chat) \equiv \sqrt{\frac{\sum_{ij} (\widehat{C}_{ij} - {C}_{ij})^2}{\sum_{ij} {C}_{ij}^2}},
\end{equation}
where $\C$ is the true covariance matrix. To compute this metric, we draw different numbers of realizations (different draws of Gaussian noise) of our toy model EoR measurement, $\textbf{x}_{\rm EoR}$, and take their ensemble average. We then compare this to the ``true' covariance, which in our simulation is set to be the empirical covariance after a large number ($500$) of realizations. As shown in Figure \ref{fig:toy_sigloss16}, we perform this computation for a range of total independent ensemble realizations (horizontal axis) and the number of independent samples in the data following fringe-rate filtering (different colors). With more independent time samples (i.e. more realizations) in the data, one converges to the true fringe-rate filtered covariance more quickly.

@@

%Just as important as the eigenvalues are the eigenvectors of our empirical covariances. 
The overall convergence of the covariance is important, but it is an important fact that, in general, different eigenvectors converge to their true forms at different rates. This is illustrated by Figure \ref{fig:toy_sigloss17}, which shows the convergence of eigenvectors in an empirical estimate of a covariance matrix whose true form is a diagonal matrix with eigenvalues spanning four orders of magnitude. We use as a convergence metric $\varepsilon(\widehat{\textbf{v}})$ for the empirical eigenvectors $\widehat{\textbf{v}}$, defined as:
\begin{equation}
\label{eq:converge_eig}
\varepsilon (\widehat{\textbf{v}}) \equiv \sqrt{\sum_{i}^{N_{f}}|\textbf{v}-\widehat{\textbf{v}}|_{i}^2},
\end{equation}
where $N_{f}$ is the number of frequencies ($20$) in the mock data. The eigenmode convergence curves  in Figure \ref{fig:toy_sigloss17} are ranked ordered by eigenvalue, such that ``Eigenmode \#0" illustrates the convergence of the eigenvector with the largest eigenvalue, ``Eigenmode \#1" for the second largest eigenvalue, and so on. One sees that the highest-valued eigenmodes converge to their true eigenvectors more quickly. With only a small number of realizations, these empirically estimated modes already retain little correlation with the specific realizations of data that were used to form the empirical covariance. As we will see in the next section, using only these eigenmodes, which are less coupled to data realizations, minimizes signal loss. In contrast, the lower-valued eigenmodes retain more memory of the data realizations, which leads to correlations that induce signal loss. 
% JEA - this is redundant
%Said differently, a steep covariance eigenspectrum can be especially dangerous because the eigenmodes that converge the slowest are typically also EoR-dominated and are therefore susceptible to the most loss.  %\acl{Think about whether we need to do the Fourier space version as well.}

@@

% JEA - death to eigenspectra!
%\begin{figure}
%	\centering
%	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.31\textwidth]{plots/toy_sigloss15.pdf}
%	\caption{Eigenspectrum of $\widehat{\textbf{C}}_{\rm FG+EoR}$, in the case of no fringe-rate filtering (solid green) and with fringe-rate filtering (dashed green). The dashed, steep eigenspectrum has a greater risk of signal loss because its lowest-valued eigenmodes are both more strongly coupled to the data (than those of the solid eigenspectrum) and, in this example, are also EoR-dominated.}
%	\label{fig:toy_sigloss15}
%\end{figure}

@@

Using our toy model, we have seen that a sensitivity-driven analysis technique like fringe-rate filtering has trade-offs of signal 
loss and noisier estimates when using data-estimated covariance matrices. Longer integrations increase sensitivity but reduce 
the number of independent samples, resulting in 
eigenmodes correlated with the data
% JEA - it's not that they are wrong, but that they are correlated
%poorly characterized, steep eigenspectra 
that can overfit signal greatly. We 
note that a fringe-rate filter does have a range of benefits, many described in \citet{parsons_et_al2016}, so it can still be 
advantageous to use one despite the trade-offs.

@@

In Section \ref{sec:toymodel} we showed one example of how altering $\widehat{\textbf{C}}$ can 
make the difference between nearly zero and some signal loss. We will now use our toy model to describe several other ways to tailor $\widehat{\textbf{C}}$ 
in order to minimize signal loss. We choose four independent regularization methods to highlight in this section, which have 
been chosen due to their simplicity in implementation and straightforward interpretations. We illustrate the resulting power 
spectra and eigenspectra for the different cases in Figures \ref{fig:toy_sigloss8}.
% and \ref{fig:toy_sigloss14}. 
These examples are not meant to be taken as suggested analysis methods but rather as illustrative cases.

As a first test, we model the covariance matrix of EoR as a proof of concept that if perfect models are known, signal loss can be 
avoided. We know that our simulated EoR signal should have a covariance matrix that mimics the identity matrix, with its 
variance encoded along the diagonal. We model $\textbf{C}_{\rm EoR}$ as such (i.e. the identity), instead of computing it based on $\textbf{x}
_{\rm EoR}$ itself. Next, we add $\textbf{C}_{\rm EoR} + \widehat{\textbf{C}}_{\rm FG}$ (where $\widehat{\textbf{C}}_{\rm FG} = \langle\textbf{x}_{\rm FG}
\textbf{x}_{\rm FG}^{\dagger}\rangle_{t}$) to obtain a final $\widehat{\textbf{C}}$ to use in weighting. In Figure \ref{fig:toy_sigloss8} (upper 
left), we see that there is negligible signal loss. This is because by modeling $\textbf{C}_{\rm EoR}$, we avoid over-fitting EoR fluctuations in the data that our model doesn't know about (but an empirically derived $\widehat{\textbf{C}}_{\rm EoR}$ would). 
%This is also shown by comparing the (steeper) green and (flatter) red curves in Figure \ref{fig:toy_sigloss14}. 
In practice such a weighting option is not feasible, as it is difficult to model $\textbf{C}_{\rm EoR}$, and $\widehat{\textbf{C}}_{\rm FG}$ is unknown because we don't know how to separate out the foregrounds from the EoR in our data.

@@

%\begin{figure}
%	\centering
%	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss14.pdf}
%	\caption{We compare the eigenspectrum of an empirically calculated $\widehat{\textbf{C}}$ (green) to that of four alternate 
%weighting options, including modeling the covariance matrix of EoR (red), regularizing $\widehat{\textbf{C}}$ by adding an identity 
%matrix to it (blue), using only the first three eigenmodes of $\widehat{\textbf{C}}$ (yellow), and multiplying an identity matrix with $
%\widehat{\textbf{C}}$ (magenta). All eigenspectra (except the green) are relatively flat and don't exhibit signal loss. All were 
%computed for the `fringe-rate filtered' (time-averaged) toy model case presented in Section \ref{sec:toymodel_frf}.}
%	\label{fig:toy_sigloss14}
%\end{figure}
%
The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\widehat{\textbf{C}} \equiv 
\widehat{\textbf{C}} + \gamma\textbf{I}$, where $\gamma = 5$ (an arbitrary strength 
of $\textbf{I}$ for the purpose of this toy model). By adding the identity matrix, element-wise, we are weighting the diagonal 
elements of the estimated covariance matrix more heavily than those off-diagonal. Since the identity component does not know anything about the data realization, it alters the covariance to be less coupled to the data. Although there is negligible signal loss using this regularization, the small green peak at the third $k$-mode represents residual foregrounds that still exist since the shapes encoded in the off-diagonal frequency correlations of the covariance matrix were deemed not as prominent as the diagonal elements using this weighting scheme.
